---
title: "Assignment 13 - Shiny and Maps (4 pts)"
author: "Group 2, Akhil Havaldar (ash2sfp)"
date: "Due Date: 11:59pm, Apr 4"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
runtime: shiny    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
library(tidyverse)
library(shiny)
library(shinythemes)
```

# Group Homework 

- You will work with your group to complete this assignment.

- Upload your Shiny app on [ShinyApps.io](https://www.shinyapps.io).

- Submit your group's shared `.Rmd` AND your URL (https://XXXXX.shinyapps.io/XXXXX) on Collab.

- Confirm this with the following comment included in your submission text box: **"Honor Pledge: I have recreated my group submission using using the tools I have installed on my own computer"**
    
- Name the files with a group name and YOUR name for your submission.

- One of the group members will **present R codes and plots for Parts 1 and 2** in class on **Apr. 4 (Mon)**. Please e-mail the instructor if you're a presenter by 11:59pm, Apr. 3. Also, if you're a presenter, please bring your laptop so that you can share your screen on zoom for the presentation.

# Part 1

- Get the latest data from from [https://covid19.who.int/info?openIndex=2](https://covid19.who.int/info?openIndex=2).
      
   - The file should likely be named "WHO-COVID-19-global-table-data.csv"
      
   - Don't use the data for the previous group assignments. It is not the most recent data. 

- Choose three WHO regions of interest. Note that we have 6 main regions: Africa, Americas, Eastern Mediterranean, Europe, South-East Asia, and Western Pacific.

- Create a subset including 3 countries per `WHO region`. You can choose any three countries within each `WHO region` to compare the mortality rate (`mutate(rate = "Deaths - cumulative total"/"Cases - cumulative total")`). You will have 9 countries (3 countries * 3 WHO regions).

- Using `navbarPage()`, create a shiny dashboard that contains 3 tabs where each tab has a barplot of mortality rates with error bars from three countries in the selected region. For example,

```
ui <- navbarPage(title = "Mortality Rate", 
      tabPanel("Africa", ...), 
      tabPanel("Americas", "...), 
      tabPanel("Eastern Mediterranean", ...))
```      
- Create separate `tabPanel()` for each level of`WHO region` variable.

- Under each tab, create a bar plot (with error bars) from three countries in the selected region with a controller for including/excluding errors bars. The default plot has no error bars.

- Use different control types under each tab. Check available control types from [Widget gallery](https://shiny.rstudio.com/gallery/widget-gallery.html).
  
- Change a theme from the `shinythemes` package. See options from https://rstudio.github.io/shinythemes/.   

- Tips for  `renderPlot()`:

```
    output$<id> = renderPlot({
     
     p1 <-  ...your ggplot...
     print(p1)
      
     if (input$<id>) { # modify this line
     p1_e <- p1 + geom_errorbar(aes(ymin=rate-1.96*SE, ymax=rate+1.96*SE), width=.2)  # add error bars
     print(p1_e)
     }}) 
```

- Example: https://youmi.shinyapps.io/A13part1/

```{r}
library(dplyr)

covids <- read_csv("WHO-COVID-19-global-table-data.csv")
focus <- c("South Africa", "Ethiopia", "Kenya", 
                   "Haiti", "Brazil", "Peru",
                   "France", "Turkey", "Italy")
c <- covids %>% filter(Name %in% focus) %>% 
  mutate(rate = `Deaths - cumulative total`/`Cases - cumulative total`, SE = sqrt(rate*(1-rate)/`Cases - cumulative total`))


ui <- navbarPage(title = "Mortality Rate", theme = shinytheme("superhero"),
                 
      tabPanel("Africa", checkboxInput("checkbox", label = "Error Bars", value = F), 
               plotOutput("africa_plot")),
      
      tabPanel("Americas", radioButtons("radio", label = h3("Error Bars"),
    choices = list("Error Bars" = T, "No Error Bars" = F), selected = F), plotOutput("americas_plot")), 
    
      tabPanel("Europe", selectInput("select", label = h3("Error Bars"), 
    choices = list("Error Bars" = T, "No Error Bars" = F), selected = F), plotOutput("europe_plot")))


server <- function(input, output) {
  
  output$africa_plot <- renderPlot({
    p1 <- ggplot(c[c$`WHO Region`=='Africa',], aes(x=rate, y=Name)) + geom_col() + theme_bw() + xlab('Mortality Rate') + coord_flip()
    
    print(p1)
    
    if (input$checkbox) {
      p1e <- p1 + geom_errorbar(aes(xmin=rate-1.96*SE, xmax=rate+1.96*SE), width=.2)
      print(p1e)
    }
  })
  
  output$americas_plot <- renderPlot({
    p2 <- ggplot(c[c$`WHO Region`=='Americas',], aes(x=rate, y=Name)) + geom_col() + theme_bw() + xlab('Mortality Rate') + coord_flip()
    
    print(p2)
    
    if (input$radio) {
      p2e <- p2 + geom_errorbar(aes(xmin=rate-1.96*SE, xmax=rate+1.96*SE), width=.2)
      print(p2e)
    }
  })
  
  output$europe_plot <- renderPlot({
    p3 <- ggplot(c[c$`WHO Region`=='Europe',], aes(x=rate, y=Name)) + geom_col() + theme_bw() + xlab('Mortality Rate') + coord_flip()
    
    print(p3)
    
    if (input$select) {
      p3e <- p3 + geom_errorbar(aes(xmin=rate-1.96*SE, xmax=rate+1.96*SE), width=.2)
      print(p3e)
    }
  })
  
}

shinyApp(ui = ui, server = server, options = list(height = 600)) 
```

# Part 2

- Create a choropleth US map like the Trump Impeachment example. You need to add any state-related variable (e.g., the number of universities per state, the crime rate per state) onto the US map.

- Use a dataset that is available in data repositories (e.g., [kaggle](https://www.kaggle.com/datasets)). But don't use the built-in datasets or datasets from packages. You need to find a state-related variable online. 

- Briefly describe the dataset you're using for this assignment (e.g., means to access data, context, sample, variables, etc...)
  
   + *describe your data.*
   
- Create a choropleth map that displays data on a map by shading states with different colors.

- You can use any map-related functions we've learned in class or any other available functions for creating maps. Your map can be a static map or an interactive map.

Data
- This dataset can be found on kaggle (https://www.kaggle.com/datasets/mathchi/violent-crime-rates-by-us-state). The data gives rates of arrests per 100,000 for murder, assault, and rape for all 50 states. This data can be well displayed oa  chloropleth map, which is the main reason for choosing this data set. 


```{r, warning=FALSE, message=FALSE}
library(ggmap)

# Insert your API key to get your google map for this session
register_google(key = "")

register_google()
library(readr)
library(ggplot2)

crime <- read_csv("US_violent_crime.csv")
names(crime)[1] <- "region"
crime$region <- tolower(crime$region)
states <- map_data("state")

map <- merge(crime, states, by="region")


ggplot(map) + 
  geom_polygon(aes(x=long,y=lat, group=group, fill=Murder)) + coord_map() + theme_bw()+scale_fill_gradient2(low="blue", mid='purple', high="red")

```


- From this map, we can see which states have the highest murder rates in the US. The more red, the higher the murder rates, and the more purple, the lower the rates are. It seems like a good amount of the southern states have high murder rates, while the midwest and more west states have lower murder rates.  
