---
title: "Assignment 15 - Final Project Proposal (2 pts)"
author: "Akhil Havaldar (ash2sfp)"
date: "Due Date: 11:59pm, Apr 5 (Tue)"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
---

<style>
h1.title {
  font-size: 30px;
}
h1 {
  font-size: 26px;
}
h2 {
  font-size: 22px;
}
h3 { 
  font-size: 18px;
}

</style>

For the describe proposal, briefly describe your topic along with both an article/report/blog and a dataset. *Submit both `.Rmd` and "knitted"`.html` files*. You **don't** need to upload your `html` file on [RPubs](https://rpubs.com).

You should not use the same topic as that for your midterm presentation. Please use a different topic for your final project. 

# Article

- https://www.afar.com/magazine/best-time-to-book-a-hotel
- This article goes over questions you should ask yourself when trying to book a hotel. In order to get the best rates at the best times you have to good timing. Hotel rates vary drastically depending on how far out you book, what time of year your traveling, what days of the week you plan on staying, etc. This article suggests you ask the following questions in order to get the best deal on your stay. First, what is the best day of the week to book? The article says Sunday for domestic travel, and Friday for international. Second, what is the cheapest day of the week to check in? Tuesday seems the best day for this. Third, how far in advance should you book? The closer to your check in date the better, since hotels tend to cut rates closer to check in dates. Fourth, what time of year should you avoid hotel stays? You should definitely avoid hotel stays during prime vacation and holiday season.

- Title: What should you look for when booking a hotel?

# Data

- https://www.kaggle.com/datasets/jessemostipak/hotel-booking-demand
- This data set contains information regarding the booking data for city hotels and resort hotels. The variables within the data set match well to the questions I am trying to answer in the article. The data set contains information on check in month, which will help determine whether or not the majority of people book hotels in the prime vacation months. The data set also contains information on lead time, which is the amount of days in advanced the hotel room was booked. This will help determine if people are taking advice of the article and booking close to their desired check in date. Lastly, I want to look at if there is a distinction in the average daily rate between resort hotels and city hotels.     


- Print the first 6 rows using function `head()`.

```{r}
library(readr)
hotel <- read_csv("hotel_bookings.csv")
head(hotel)
```


# Data Validation

- Check the content of the attributes, traits, features, rows, and/or other properties specific to your dataset.

   + Is the data type correct for this field?

   + Is the value within the valid range or part of a domain or enumerated list?

   + Check for duplicates, for example of a unique key.

   + Check for nulls. Are there mandatory values, or are null / empty values allowed? Are the null types consistent (NaN, infinity, empty strings, etc.)?

- Tips: for duplicates, use function `unique()`. For missing values, use function `is.na()`. To delete all rows with a missing, use function `na.omit()`. Feel free to use other functions to check duplicates and missing values.

- Since the data set contains almost 120,000 observations, I am removing all NAs in the data set. This brings the number of observations down from 119,390 to 119,386 (only 4 removed).

```{r}
hotel <- na.omit(hotel)
```


# Plot 

- You need to create at least one plot using `shiny`.

- Describe your key plot(s) that you like to create in words.

- I plan to create a few shiny plots. One being a graph of average daily rates with a radio button to select between city and resort hotels. I also want to make a bar/line plot for both lead time and the arrival month. Both of which can have radio buttons to change between resort and city hotels. 


# Post on Discussion Board

- Check your presentation date. 

- Create a post on the discussion board with your presentation title as the post title by 11:59 pm, Apr 5. You'll edit your post on the discussion board to include your presentation files (Rmd file and your shiny-app URL).

- You should upload your presentation files (Rmd file and your shiny-app URL) and dataset you used on the discussion board by the day before your presentation day. For example, if you're a presenter on Apr 11, upload your presentation files by 11:59 pm, Apr 10.  


